---
title: "Case_Study_I_Beer_DistributionBy_We'veGotData"
author: "Jayson Baker, Jeremy Simpson, and William Hinton"
date: "6/23/2019"
output: 
  html_document: 
    keep_md: true
---
# Introduction 
This report provides an analysis of craft beers and Brewery distributors of craft beers throughout the US.  Your team has expressed interest in understanding the location of industry competitors, market saturation per state, along with assistance in determining the composition of your next beer product, according to alcohol content (ABV) and bitterness (IBU). The prepared analysis will accommodate you in addressing your needs. Based on this analysis along with follow-up discussions with your team and other relevant business partners, we feel that you will be positioned to make a strong business decision for your organization.

```{r include=FALSE}
library(tidyverse)
library(sqldf)

setwd("F:/SMU/DS6306/Case_Study/DS6306_CS1")
getwd()
```

## Specifications
>*Here we display the number of Breweries by each state. This will allow you to know who your competitors are and what state has the most Breweries which should assist with favorable location identification. We suggest also factoring in the composition of ABV and IBU for beer products in those states before choosing a location.*

```{r echo=TRUE}
# Read the data files
b_beers <- read.csv("./Raw_Data/Beers.csv")
str(b_beers)

a_breweries <- read.csv("./Raw_Data/Breweries.csv")
str(a_breweries)

```

>*Here are counts of Breweries listed by each State. In total we observed 558 breweries with the largest concentration of breweries located in Colorado. *

```{r echo=TRUE}
a1_breweries <- data.frame(sqldf("SELECT State, COUNT(*) as BrewCount FROM a_breweries GROUP BY State ORDER BY COUNT(*) DESC"))  # count of breweries by state
a1_breweries
```

>*We merged the two data files and listed the top 6 records and the last 6 records in the merged file to examine the data post merge.*

```{r echo=TRUE}
b1_beers <- data.frame(sqldf("SELECT * FROM a_breweries a INNER JOIN b_beers b ON a.Brew_ID = b.Brewery_ID"))  # merge the two files

head(b1_beers,6)  # list the first 6 obs in merged file
tail(b1_beers,6)  # list the last 6 obs in the merged file

```

>*We identify the number of "NA's" in each columns so that we can cleanup the data and make it more readable.* 
>*We detected NA values primarily in the IBU and ABV column which matches the raw data. Please keep this in mind when analyzing results.*

```{r echo=TRUE}
c_na <- colSums(is.na(b1_beers))  # count the number of NAs in each column
c_na                               

```

>*Here are alcohol content and IBU values are summarized for each state using the median values of each.* 

```{r echo=TRUE}
d_median <- data.frame(sqldf("SELECT State, MEDIAN(ABV), MEDIAN(IBU) FROM b1_beers GROUP BY State"))
str(d_median)


```

#### Included below are Plots
>*Here we plot the median ABV and IBU respectively by each state to help identify which states may be more favorable and/or primed for your product positioning.*

```{r fig.width=8, fig.height=10,echo=TRUE}
# create the graph - Median ABV
ggplot(d_median, aes(x = reorder(State, MEDIAN.ABV.), y = MEDIAN.ABV., fill = -MEDIAN.ABV.)) + 
  geom_bar(stat = "identity", position="dodge",  na.rm = TRUE) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + 
  ggtitle("Median Alcohol Content by Volume (ABV) by State") + theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x = "State (U.S.)", y = "Alcohol Content by Volume (ABV)") +
  coord_flip() # flips the chart
  
# create the graph - Median IBU
ggplot(d_median, aes(x = reorder(State, MEDIAN.IBU.), y = MEDIAN.IBU., fill = -MEDIAN.IBU.)) + 
  geom_bar(stat = "identity", position="dodge",  na.rm = TRUE) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + 
  ggtitle("Median International Business Unit (IBU) by State") + theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x = "State (U.S.)", y = "International Business Unit (IBU)") +
  coord_flip() # flips the chart
```

### Alcohol Content
>*This section will list and rank maximum alcoholic content for ABV by state in descending order and the most bitter IBU beer by state in descending order.*
>*The top two states listed below one for each category (ABV, IBU), (Colorado for max ABV) and (Oregon for bitter IBU).* 

```{r echo=TRUE}
e <- data.frame(sqldf("SELECT State, MAX(ABV) FROM b1_beers GROUP BY State ORDER BY ABV DESC"))  # rank order by state and ABV to check
e <- data.frame(sqldf("SELECT State, MAX(ABV) FROM b1_beers where ABV = (SELECT MAX(ABV) FROM b1_beers) GROUP BY State"))  # state with the highest ABV
head(e)

f <- data.frame(sqldf("SELECT State, MAX(IBU) FROM b1_beers GROUP BY State ORDER BY IBU DESC"))  # rank order by state and IBU to check
f <- data.frame(sqldf("SELECT State, MAX(IBU) FROM b1_beers where IBU = (SELECT MAX(IBU) FROM b1_beers) GROUP BY State"))  # state with the highest IBU
head(f)
```

### Summary by columns for ABV and IBU
>*Here we provide summary statistics for the ABV and IBU values to help provide a perspective on the specific ranges and statistics.*

```{r echo=TRUE}

summary(b1_beers$ABV)  # summary stats for ABV
summary(b1_beers$IBU)  # summary stats for IBU
```

### Included below are Plots
>*Here we plot the ABV and IBU for each state in an attempt to discern if there is a relationship between ABV and IBU.*
>*Note: there does appear to be a linear relationship between ABV and IBU; we've added in ounces as a dimension for each point as well.*
>*With regards to this analysis, when determine where to position your product, please keep in mind that the higher the ABV, the higher the IBU which could also factor into location preferences as Colorado has the highest ABV but Oregon has the highest IBU.*

```{r echo=TRUE}
ggplot(b1_beers, aes(x=ABV, y=IBU, color = factor(b1_beers$Ounces))) + 
  geom_point(na.rm = TRUE) +
  ggtitle("Alcohol Content by Volume (ABV) vs. International Bitterness Unit (IBU)") + theme(plot.title = element_text(hjust = 0.5)) + 
  labs(x = "Alcohol Content by Volume (ABV)", y = "International Bitterness Unit (IBU)", color = "Ounces")
```

### Summary

Based on the problem statement your team initially asked us to examine:

+ Competitor location and market saturation in each state.
+ Assistance in determining the composition of your next beer product (in terms of alcohol by volume (ABV) and bitterness (IBU))* 

The following are our recommendations based the requested analysis:

+ Avoid constructing your new brewery in Colorado.
Our analysis indicates that the state is heavily saturated with breweries which would mean stiff competition. Our suggestion, is that you give strong consideration to exploring the markets in the states of Texas and Pennsylvania. These states have a mid-level concentration of breweries and appear to have some relative room for additional market growth and competition.
+ When developing your next beer product, please consider that our analysis strongly  implies that a higher ABV will affect the bitterness of the beer. If bitterness is a key trait for your prospective product, it is probable that this can be achieved with  a high ABV concentration. If you prefer that your product is less bitter, we suggest limiting the ABV to reasonable amounts.
+ Additionally we suggest considering locations of interest when determining the bitterness level of your proposed product. Our data shows that Oregon is the state with the highest concentration of bitter beers, yet the number of breweries is well below the number present in Colorado. This is either means there is a high demand for bitter beer, or is a signal of a weak market for great tasting beer in Oregon, that has the potential to be further explored.  At minimum we suggest further exploration.